Sperm do not evolve to collaborate in female meiotic drive.

### 2014-04-27 writeup_and_notes/male_dependent_drive.tex
added heterozygous bistabiltiy analysis and editeed through single locus sperm effect

\section*{Introduction}

Despite the apparent unity of the organism, alleles can
gain an evolutionary advantage at a cost to  individual fitness
\citep{Burt2006}, often by exploiting meiosis and gametogenesis.
Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
An allele that biases female meiosis in its favor (i.e. a meiotic driver), may increase in frequency even if this driver entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
Meiotic drivers observed in both plants
\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
conflict -- the selfish benefits of drive and the associated
pleiotropic fitness costs sustain a balanced polymorphism
\citep{Prout1973}, 
and often generate ongoing evolutionary escalations of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
The threat of meiotic drive to organismal fitness is potentially so
severe that many basic properties of
meiosis and o\"{o}genesis, including the initial genome doubling in
meiosis I \citep{Haig1991}, arrested female meiosis
\citep{Mira1998}, centromere machinery \citep{Malik2002a,Malik2009,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}.

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is competed only after fertilization \citep{Masui_book}, 
creating ample opportunity for interaction between the sperm and female meiotic machinery. 
Therefore, an allele in sperm that facilitates meiotic drive by a genetically equivalent allele in female meiosis %(a  heteromorphic dyad), 
     could presumably rapidly spread through the population. 
At first sight, female meiosis appears primed for exploitation by selfish sperm systems.  
Why then is the requirement of fertilization to complete female meiosis so ubiquitous?

It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animal taxa, 
	female meiosis is complete upon ovulation (that is fertilization is not required for female meiosis \citep[see table 1 of ref 
][]{Masui_book}).

It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
In \emph{C. elegans} premature deployment of the aster, a vital component of mitotic
machinery, provided by the sperm disrupts MII meiotic segregation
in the egg, leading to a triploid zygote \citep{McNally2012}. 
Thus, there is direct mechanistic evidence that sperm can influence female meiosis.
 Additionally, genetic evidence from the \emph{In} and \emph{Om}  drive systems in mice suggests that the 
 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
 While ruling out the alternative hypothesis of early selection on zygotes \citep[pages
52-54][]{Burt2006} in these cases is challenging, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
% Both of these female drive systems operate by distorting the second meiotic division, 
% and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
%While such genetic evidence is suggestive, we note the difficulty in
%ruling out the alternative hypothesis of early selection on zygotes \citep[pages
%52-54][]{Burt2006}.

Here, we develop population genetic models exploring the evolution of sperm influence on female meiotic drive. 
%In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
These models show that sperm modifiers of female meiotic drive are
unlikely to create a sustained conflict, and therefore 
female meiosis will have little incentive to evolve resistance to them.
In fact, the interests of sperm and egg  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the In locus \citep{Pomiankowski1993}).
Therefore, there is little selective benefit to females in preventing sperm to influence female meioses,
	indeed features of meiosis may evolve that facilitate the interaction of sperm with female meiosis.

@@

In the standard single-locus, biallelic model of female meiotic drive, the driving allele is transmitted to the egg  in heterozygotes with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
To depict a case of a self-promoting meiotic
driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
\ref{Eggsperm_2_allele_cartoon}) \footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
	so $d$ is bounded to be $<0.75$? \citep[see ][ for discussion]{Buckler1999}. Could also put this in
        caption of fig.}.
We then identify the conditions under which this self propagating driver can spread, and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
\yb{EQUATIONS A1-A4}

We first consider a driving allele neutral in heterozygotes, 
	but deleterious in homozygotes (whose relative fitness is
        $1-s$). This model is laid out in eqn. \ref{eqn.A1}-\ref{eqn.A3}.
In this model, a standard meiotic driver can always invade because 
	when rare it occurs predominantly in heterozygotes and it therefore drives without a fitness cost. 
However, a severe homozygous fitness cost prevents  
	drivers from fixing  \citep{Prout1973}, 
	providing an opportunity for the evolution of
	drive suppressors (CITE - citep{XX}), corresponding well to 
	empirical examples of female meiotic drive.

In contrast to a traditional driver, which drives but pays effectively
no fitness cost as a rare deleterious recessive, 
	a self-promoting driver creates homozygotes
        when it drives, and therefore there is a drive-associated fitness cost even when rare. 
The self-promoting driver must overcome this homozygous fitness cost simply to spread when rare.
Therefore with a recessive fitness cost,
	the conditions allowing the invasion of a self-promoting driver
 	are far more restrictive than those for a standard meiotic driver (see Figure \ref{Invasion_space}).
\yb{When fully recessive, a self-promoting driver can only invade when $s$, the selection against the drive homozygote, is less than $\frac{2 d - 1}{4 d}$ (The tan region in Figure  \ref{Invasion_space}).}

When a deleterious recessive self-promoting driver does spread, 
	it spends much time at low frequency, because when rare it is unlikely to meet a complimentary sperm. 
However, once relatively common, a self-promoting driver spreads rapidly to fixation due to its
	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
This lack of a balanced polymorphism ( Figure \ref{Invasion_space}) precludes the evolution of an
allele that suppresses this form of meiotic drive in females.
\yb{In fact, while a rare self promoting driving can only invade if $s<\frac{2 d - 1}{4 d}$, a common self-promoting driver can fix over a broader range of parameter space (i.e. if $s<\frac{2d-1}{2}$). 
This system therefore displays bistability, when $\frac{2 d - 1}{4 d}<s<\frac{2d-1}{2}$, the driver will deterministically fix when introduced at a frequency greater than $\frac{1-2d+4ds}{2s(2d-1)}$, but will be lost otherwise.}

Inclusion of a heterozygous fitness cost (e.g. additive fitness cost) further constrains the evolution of a self-promoting driver. 
While a standard female driver invades from low frequency when the transmission advantage outweighs the fitness cost (XXX),
	a rare self-promoter pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
This frequency-independent fitness cost, and positive frequency dependent drive generates a bistable system. 
Below a cutoff frequency the self-promoter driver deterministically decreases in frequency 
	because it pays the  heterozygous fitness cost but drives
        ineffectively (see Figure S\ref{bistable_additive}). 
Above this frequency, the allele increases deterministically because the commonness of complementary sperm 
	make drive effective enough to overcome the associated fitness cost.
This bistability prevents self-promoting drivers from invading 	
	reasonably sized populations, and assures that if they do invade they will rapidly fix.
Our model therefore predicts that self-promoting drivers will not be observed as polymorphisms in natural populations.

@@

\begin{figure}
\includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
\caption{Invasion analysis for a female meiotic drive allele with
  recessive costs. First consider a simple female meiotic
  drive system, where the allele is transmitted a fraction $d>1/2$ of
  meioses in female heterozygotes, inspective
  of the allele carried by the fertilizing sperm. Such an allele can
  invade over all of the parameter space, and will be maintained as a
  protected polymorphism in the white region, and can fix in the
  population in all of the parameter space colored green,
  and will fix  or orange. Now consider an allele that is fairly transmitted through
  female meiosis when fertilized by a sperm carrying a different
  allele, but a fraction $d>1/2$ when a heterozygote egg is fertilized
  by a sperm carrying the allele (as depicted in Figure
  \ref{Eggsperm_2_allele_cartoon}). Such an allele can only invade in
  the region of parameter space colored orange, and it fixes over all
  of this parameter region. 
 In the inset we show trajectories of sperm-dependent female drive
 alleles that can invade the population, each allele has $s =0.1$ against the
 homozygotes and $d =0.6$, $0.7$, $0.8$ shown in yellow, blue, and red. 
 \gc{Likely make this Figure 1B.} } \label{Invasion_space}
\end{figure}

@@

Sexual reproduction is a high-stakes event that determines what gets transmitted to the next generation.  
As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
	can succeed evolutionarily, even if they lower the 
        fitness of the population, 
	the mother whose offspring they will inhabit, 
	or the organism in which they reside. 
This generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
	and conflicts between alleles that are overtransmitted in meiosis and the organisms they injure while doing so. 
Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}.

@@

Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
By contrast a sperm promotor of female drive is only effective when matched with a heterozgyote female -- meaning that when rare this allele rarely enhances female drive. 
Even worse, when it does so it will preferentially find itself in a low fitness drive homozygote. 
Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do maintain a polymorphisms and 
	provide evolution  with time and opportunity to further minimize drive.
This is because such drive suppressing alleles reduced their chances
of forming low fitness homozygotes. 
In fact, an allele in sperm which causes the alternate allele to drive can be favored (FIG), a counterintuitive result 
	consistent with the tenuous evidence that the non-driving
        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
        sperm (CITE cite{agulnick, Pomiankowski1993}).   
More generally, alleles that act through sperm to
 reduce the opportunity of female meiotic drive regardless of linkage or phase.

@@

its change in frequency across generations $\Delta f_B$ is
\begin{equation}
\frac{f_A f_B^2}{\bar{w}} \left( [1-F][ d(1-2 f_A s) - f_B s-1/2] - s  F \right)
\end{equation}
where $F$ is the deviation from Hardy-Weinberg expectations, and $f_A$ is the frequency of the non-driving allele. 
For analytical solutions presented above, we assume that $F$ 
}

\paragraph{cost of driver in heterozygotes}
We also explored a range of other costs to the drive alleles other than a
simple recessive cost, for example an additive cost. We modified the recursions given in eqn
\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
copy of the $B$ allele. A none recessive cost to the drive allele can
cause bistability of the drive allele. The B allele is unable to enter
the population when rare, this results from the fact that
when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
cost in heterozygous individuals.  By contrast, when the B allele is common ($x_B
\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
become more common allowing drive to occur this can allow the $B$
allele to spread when common. In Supplementary Figure
\ref{bistable_additive} we show the unstable equilibrium frequency
above which the $B$ allele can invade. This equilibrium frequency was found by a simple
parameter search.  \gc{wonder if a bit of work could find this}

>>>>

\section*{Introduction}
Despite the apparent unity of the organism, alleles can
gain an evolutionary advantage at a cost to  individual fitness
\citep{Burt2006}, often by exploiting meiosis and gametogenesis.
Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
An allele that biases female meiosis in its favor (i.e. a meiotic driver), may increase in frequency even if this driver entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
Meiotic drivers observed in both plants
\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c} highlight this
conflict -- the selfish benefits of drive and the associated
pleiotropic fitness costs sustain a balanced polymorphism
\citep{Prout1973}, 
and often generate ongoing evolutionary escalations of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
The threat of meiotic drive to organismal fitness is potentially so
severe that many basic properties of
meiosis and o\"{o}genesis, including the initial genome doubling in
meiosis I \citep{Haig1991}, arrested female meiosis
\citep{Mira1998}, centromere machinery \citep{Malik2002a,Malik2009,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}.

It is therefore somewhat surprising that despite the intense evolutionary pressure on female meiosis to prevent meiotic drive, 
it is potentially open to sabotage by a virtual stranger -- a haploid sperm genome.
That is, in many animal species, female meiosis is competed only after fertilization \citep{Masui_book}, 
creating ample opportunity for interaction between the sperm and female meiotic machinery. 
Therefore, a `green-bearded' \citep{Gardner2010} sperm-acting allele that recognizes and facilitates the meiotic drive of a genetically equivalent allele in female meiosis %(a  heteromorphic dyad), 
     could presumably rapidly spread through a population. 
At first sight, female meiosis appears primed for exploitation by selfish sperm systems.  
Why then is the requirement of fertilization to complete female meiosis so ubiquitous?

It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animals, 
	female meiosis is complete upon ovulation (that is fertilization is not required for female meiosis \citep[see table 1 of ref 
][]{Masui_book}).

It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
In \emph{C. elegans}, premature deployment of the aster, a vital component of mitotic
machinery provided by the sperm, disrupts MII meiotic segregation
in the egg, leading to a triploid zygote \citep{McNally2012}. 
Thus, there is direct mechanistic evidence that sperm can influence female meiosis.
 Additionally, genetic evidence from the \emph{In} and \emph{Om}  drive systems in mice suggests that the 
 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
 While ruling out the alternative hypothesis of early selection on zygotes \citep[pages
52-54][]{Burt2006} in these cases is challenging, it appears that the extent to which \emph{In} and \emph{Om} distort the second female meiotic division partially depends on the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
% Both of these female drive systems operate by distorting the second meiotic division, 
% and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
%While such genetic evidence is suggestive, we note the difficulty in
%ruling out the alternative hypothesis of early selection on zygotes \citep[pages
%52-54][]{Burt2006}.

We develop population genetic models exploring the evolution of sperm influence on female meiotic drive, focussing on models in which `self-promoting' alleles in sperm facilitate drive of like genotypes during gametogenesis in heterozygous females.  
%In this article we explore through simple population genetic models the consequences of alleles that influence the outcome of female meiosis.
These models show that sperm modifiers of female meiotic drive are
unlikely to create a sustained conflict, and therefore 
female meiosis will have little incentive to evolve resistance to them.
In fact, the interests of sperm and egg  genomes' are often aligned, as both are invested in the fate of the resultant zygote (as was speculated for the \emph{In} locus \citep{Pomiankowski1993}).
Therefore, there is little selective benefit to females in preventing sperm to influence female meioses,
	indeed features of meiosis may evolve that facilitate the interaction of sperm with female meiosis.

@@

In the standard single-locus, biallelic model of female meiotic drive, the driving allele is transmitted to the egg  in heterozygotes with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
To depict a case of a self-promoting meiotic
driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
\ref{Eggsperm_2_allele_cartoon}). 
%\footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
%	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
%	so $d$ is bounded to be $<0.75$? \citep[see ][ for discussion]{Buckler1999}. Could also put this in
  %      caption of fig.}.
We then identify the conditions allowing the spread of this self-promoting driver, 
	and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
%\yb{ We present a general framework for this problem in equations A1-A4, and evaluate specialized cases below.}

For comparative purposes, we present the standard drive model in equations \ref{eqn.A1}-\ref{eqn.A3}, 
	and refer readers to the previous literature on this model (\citep[e.g. ][]{Prout1973,Ubeda2004} for additional results). 
In this case, assuming the driving allele is deleterious but fully recessive 
	(i.e. homozygous fitness equals $w_{BB}=1-s$ and other genotypic fitnesses equal $w_{AA}=w_{AB}=1$), 
	it always invades because,  when rare, it occurs predominantly in heterozygotes and therefore drives without a fitness cost. 
However, when $s$ is large drivers cannot fix \citep{Prout1973}, 
	providing an opportunity for the evolution of
	drive suppressors, corresponding well to 
	empirical examples of female meiotic drive \citep[reviewed in ][]{Burt2006}.

In contrast to a traditional driver, which drives but pays effectively
	no fitness cost when rare, 
	a self-promoting driver specifically creates homozygotes by driving. 
%, and therefore there is a drive-associated fitness cost even when rare. 
Thus it must overcome a drive-associated homozygous fitness cost simply to spread when rare,
	and the conditions allowing the invasion of a self-promoting driver
 	are far more restrictive than those for a standard meiotic driver.
When rare, a fully recessive, self-promoting driver can only invade when $s$ 
	%, the selection against the drive homozygote, 
	is less than $\frac{2 d - 1}{4 d}$. 
This analytical result, derived  from Equation \eqref{deltadriver} assuming Hardy-Weinberg, 
	closely matches results obtained by exact numerical iteration (Figure \ref{Invasion_space}).

When a self-promoting driver does spread, 
	it spends much time at low frequency, 
	because the paucity of complimentary sperm compromises its ability to drive. 
However, once relatively common, it  rapidly achieves fixation due to its
	positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}).  
This lack of a balanced polymorphism (Figure \ref{Invasion_space}) precludes the evolution of an
	allele that suppresses this form of meiotic drive in females.

The positive frequency dependence of a self-promoting driver can induce bistability in its spread. 
When rare, a  recessive self-promoting driver can spread only if $s < \frac{2 d - 1}{4 d}$;
	however, once common such a driver can fix over a broader range of parameter space (i.e. if $s<\frac{2d-1}{2}$). 
Therefore, when $\frac{2 d - 1}{4 d}<s<\frac{2d-1}{2}$, 
	a recessive self-promoting driver will deterministically fix if drift, mutation, 
	or migration pressure bring its frequency above $\frac{1-2d+4ds}{2s(2d-1)}$, 
	but it will be lost otherwise.
Inclusion of a heterozygous fitness cost (i.e. $w_{AB}=1-h_s$)
	%(e.g. additive fitness cost) 
	further constrains the evolution of a self-promoting driver. 
This case also displays bistability -- a rare self-promoting driver cannot spread; 
	however, if $s < \frac{2 d - 1+s_h(3-2d)}{2}$ it will fix deterministically if its 
	frequency is greater than some threshold, $f_B^*$ [Equation \eqref{thresh2}, and \yb{FIGURES SXX}].
%	because it pays the  heterozygous fitness cost but drives
 %      ineffectively (see Figure S\ref{bistable_additive})
%While a standard female driver invades from low frequency when the transmission advantage outweighs the fitness cost (XXX),
%	a rare self-promoter pays a fitness cost, but drives only weakly because complementary sperm are uncommon. 
%This frequency-independent fitness cost, and positive frequency dependent drive generates a bistable system. 
%Below a cutoff frequency the self-promoter driver deterministically decreases in frequency 
%	because it pays the  heterozygous fitness cost but drives
 %      ineffectively (see Figure S\ref{bistable_additive}). 
%Above this frequency, the allele increases deterministically because the commonness of complementary sperm 
%	make drive effective enough to overcome the associated fitness cost.
This bistability prevents self-promoting drivers from invading 	
	reasonably sized populations, and assures that, if they do invade, they will rapidly fix.
Our model therefore predicts that self-promoting drivers will not be observed as polymorphisms in natural populations.

@@

\begin{figure}
\includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
\caption{Invasion analysis for a female meiotic drive allele with
  recessive costs. First consider a simple female meiotic
  drive system, where the allele is transmitted a fraction $d>1/2$ of
  meioses in female heterozygotes, inspective
  of the allele carried by the fertilizing sperm (we note that the timing of fertilization relative to female meiosis places another constraint on $d$, for example, when drive takes place druing the second meitoic division
	it requires an uneven number of crossovers between the centromere and the drive locus \citep[see ][ for discussion]{Buckler1999}). 
Such an allele can invade over all of the parameter space, and will be maintained as a
  protected polymorphism in the white region, and can fix in the
  population in all of the parameter space colored green,
  and will fix  or orange. Now consider an allele that is fairly transmitted through
  female meiosis when fertilized by a sperm carrying a different
  allele, but a fraction $d>1/2$ when a heterozygote egg is fertilized
  by a sperm carrying the allele (as depicted in Figure
  \ref{Eggsperm_2_allele_cartoon}). Such an allele can only invade in
  the region of parameter space colored orange, and it fixes over all
  of this parameter region. 
 In the inset we show trajectories of sperm-dependent female drive
 alleles that can invade the population, each allele has $s =0.1$ against the
 homozygotes and $d =0.6$, $0.7$, $0.8$ shown in yellow, blue, and red. 
 \gc{Likely make this Figure 1B.} } \label{Invasion_space}
\end{figure}

@@

Sexual reproduction is a high-stakes event that determines what gets transmitted to the next generation.  
As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
	can succeed evolutionarily, even if they lower the 
        fitness of the population, 
	the mother whose offspring they will inhabit, 
	or the organism in which they reside. 
This generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
	and conflicts between alleles that are over-transmitted in meiosis and the organisms they injure while doing so. 
Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}.

@@

Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
By contrast a sperm promotor of female drive is only effective when matched with a heterozygote female -- meaning that when rare this allele rarely enhances female drive. 
Even worse, when it does so it will preferentially find itself in a low fitness drive homozygote. 
Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do maintain a polymorphisms and 
	provide evolution  with time and opportunity to further minimize drive.
This is because such drive suppressing alleles reduced their chances
of forming low fitness homozygotes. 
In fact, an allele in sperm which causes the alternate allele to drive can be favored (FIG), a counterintuitive result 
	consistent with the tenuous evidence that the non-driving
        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
        sperm (CITE cite{agulnick, Pomiankowski1993}).   
More generally, alleles that act through sperm to
 reduce the opportunity of female meiotic drive regardless of linkage or phase.

@@

its change in frequency across generations $\Delta f_B$ is
\begin{equation}
\frac{f_A f_B^2}{\bar{w}} \left( [1-F][ d(1-2 f_A s) - f_B s-1/2] - s  F \right)
\label{deltadriver}
\end{equation}
where $F$ is the deviation from Hardy-Weinberg expectations, and $f_A$ is the frequency of the non-driving allele. 
For analytical solutions presented above, we assume that $F$ equals zero. 
}

\paragraph{cost of driver in heterozygotes}
We explored a more complex model where the self-promoting driver has a cost, $s_h$ in heterozygotes (so $w_{AB} = 1-h_s$) in addition to the cost $s$, in homozygotes.
Model construction is similar to that above.
So long as there is any heterozygous fitness cost (i.e. $s_h > 0$) the self-promoting driver cannot invade when rare;
	however, as is the case without a heterozygous cost, this system is bistable.
Assuming no deviation from Hardy-Weinberg a self-promoting driver with a heterozygous fitness cost deterministically fixes if it's frequency exceeds some threshold, $f_B^*$.

\begin{equation}
	f_B^* = \frac{a_1+\sqrt{a_1^2+8 s_h a_2}}{2 a_2}
	\label{thresh2}
\end{equation}
where, $a_1=1-2 ds_h+4 d s-2 d-3 s_h$ and $a_2=2(s-s_h)(2d-1)$.

 range of other costs to the drive alleles other than a
simple recessive cost, for example an additive cost. We modified the recursions given in eqn
\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
copy of the $B$ allele. A none recessive cost to the drive allele can
cause bistability of the drive allele. The B allele is unable to enter
the population when rare, this results from the fact that
when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
cost in heterozygous individuals.  By contrast, when the B allele is common ($x_B
\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
become more common allowing drive to occur this can allow the $B$
allele to spread when common. In Supplementary Figure
\ref{bistable_additive} we show the unstable equilibrium frequency
above which the $B$ allele can invade. This equilibrium frequency was found by a simple
parameter search.  \gc{wonder if a bit of work could find this}
